Liveness-Based Garbage Collection for Lazy Languages

### 2015-04-28 paper/paper.tex
Added latest graphs

\section{Collecting more garbage can never slow things down}
\label{sec:lgc-always-better}
Since garbage collection is  effectively asynchronous to the allocator
thread, one might  worry as to how {\em  robust} our measurements are.
For example,  while LGC would,  in general, collect more  garbage than
RGC in the same heap state,  is it possible that on some programs, LGC
might do  a larger number  of collections? We  prove a result  to show
that   this  cannot   happen.    This  result   applies  to   classical
mark-and-sweep and copying garbage  collectors and, we hypothesize, to
generational collectors too.

\begin{lemma}
For the  same mutator,  a liveness-based collector  can never  do more
garbage collections than a reachability-based collector.
\end{lemma}

\begin{proof}Assume, as before,  that time is measured in terms of the number of
  cons cells allocated. Let us run  two copies of the mutator with the
  two    different   collectors    but    with   synchronous    memory
  allocations.   The   garbage    collections,   however,   happen
  asynchronously.

  To prove the lemma, it is enough to show the truth of the following 
statement: After
 every  LGC invocation,  the count of LGC invocations   is  no greater
than  RGC invocations.   The  base case  holds  since the  first
invocations of both GCs happen at the same time.  Assume the statement
to be true after $n$ invocations of LGC.  Since LGC copies a subset of
reachable cells, its heap would contain no more cells than RGC heap at
the end  of the $n$th invocation.   Thus either RGC  is invoked next before
LGC, or LGC and RGC are both invoked next at the same time.  In either
case, the statement holds after $n+1$ invocations of LGC.
\end{proof}

>>>>

\section{Collecting more garbage can never slow things down}
\label{sec:lgc-always-better}

%% Since garbage collection is  effectively asynchronous to the allocator
%% thread, one might  worry as to how {\em  robust} our measurements are.
%% For example,  while LGC would,  in general, collect more  garbage than
%% RGC in the same heap state,  is it possible that on some programs, LGC
%% might do  a larger number  of collections? We  prove a result  to show
%% that   this  cannot   happen.    This  result   applies  to   classical
%% mark-and-sweep and copying garbage  collectors and, we hypothesize, to
%% generational collectors too.

%% \begin{lemma}
%% For the  same mutator,  a liveness-based collector  can never  do more
%% garbage collections than a reachability-based collector.
%% \end{lemma}

%% \begin{proof}Assume, as before,  that time is measured in terms of the number of
%%   cons cells allocated. Let us run  two copies of the mutator with the
%%   two    different   collectors    but    with   synchronous    memory
%%   allocations.   The   garbage    collections,   however,   happen
%%   asynchronously.

%%   To prove the lemma, it is enough to show the truth of the following 
%% statement: After
%%  every  LGC invocation,  the count of LGC invocations   is  no greater
%% than  RGC invocations.   The  base case  holds  since the  first
%% invocations of both GCs happen at the same time.  Assume the statement
%% to be true after $n$ invocations of LGC.  Since LGC copies a subset of
%% reachable cells, its heap would contain no more cells than RGC heap at
%% the end  of the $n$th invocation.   Thus either RGC  is invoked next before
%% LGC, or LGC and RGC are both invoked next at the same time.  In either
%% case, the statement holds after $n+1$ invocations of LGC.
%% \end{proof}
